---
title: "DASHBOARD" 
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    social: ["menu"]
    source_code: embed
    vertical_layout: fill
    theme: united
---

This dashboard displays energy data from  [BP Statistical Review of the World 2019](https://www.bp.com/en/global/corporate/energy-economics/statistical-review-of-world-energy.html).

``` {js}
$('.navbar-inverse').removeClass('navbar-inverse').addClass('navbar-default');
```

```{r setup, include=FALSE}

library(flexdashboard) # Dashboard package
```

<style>                     
.navbar {
  background-color:#003D79;
  border-color:white;

}

.custom {
    background-color: aliceblue;
    border-color: #ffffff;
}

</style>


Primary Energy {data-icon="fa-signal"}
=======================================================================

 Column {.tabset .tabset-fade data-width=550 .custom }
-----------------------------------------------------------------------

### __World enegy consumption__ {.no-padding data-icon="fa-signal"}
```{r fig.height=3, message=FALSE, warning=FALSE}


```


### __Consumption by fuel__  {.no-padding}
```{r fig.height=3}


```


### __Contribution to primary energy growth__ {.no-padding}
```{r fig.height=3}


```


Column {data-width=300}
-----------------------------------------------------------------------

```{r include=FALSE}


```

### Growth of global primary energy consumption.

```{r}
valueBox("... %", icon = "fa-percent", color = "orange")
```

### By fuel, energy consumption growth.

```{r}
valueBox("... %", icon = "fa-percent", color = "orange")
```

### Global increase in energy demand.

```{r}
valueBox("... %", icon = "fa-percent", color = "orange")
```


Oil {data-icon="fa-signal"} 
========================================================================

Column {.tabset .tabset-fade data-width=550 .custom }
-----------------------------------------------------------------------

### __Map of Oil Reserves by country in 2018__ {.custom }
```{r fig.height=3}

```

### __Global oil production__ {.custom }
```{r fig.height=3}


```

### __Oil proved, production & consumption__ {.no-padding .custom }
```{r fig.height=3}

  
```

### __Refinery throughput__ {.no-padding .custom }
```{r fig.height=3}


```



### __Regional refining margins__ {.no-padding .custom }
```{r fig.height=3}



```

### __Spot crude prices ($/bbl)__ {.no-padding}
```{r fig.height=3}

```


Column {data-width=300}
-----------------------------------------------------------------------


```{r include=FALSE}


```

### Annual average Dated Brent 2018.

```{r}
 
```

### Oil consumption.

```{r}

```

### Global oil production.

```{r}

```

### Refinery throughput.

```{r}
 
```


Gas {data-icon="fa-signal"}
========================================================================

Column {.tabset .tabset-fade data-width=550 .custom }
-----------------------------------------------------------------------

### __Map of Gas Reserves by country in 2018__ {.custom }
```{r fig.height=3}



```

### __Natural gas consumption__ {.no-padding}
```{r fig.height=3}




```

### __Growth in gas consumption__ {.no-padding}
```{r fig.height=3}


```

### __Global natural gas production__ {.no-padding}
```{r fig.height=3}


```

### __Growth in inter-regional natural gas trade__ {.no-padding}
```{r message=FALSE, warning=FALSE}

```

### __LNG trade movements 2018__ {.no-padding}
```{r }
 
```

### __Gas prices ($/Million Btu)__ {.no-padding}
```{r fig.height=3}
 
```

Column {data-width=300}
-----------------------------------------------------------------------


```{r include=FALSE}
 
```

### Global natural gas production.

```{r}
 
```


###  Natural gas consumption.

```{r}
 
```

### Growth in gas consumption.

```{r}
 
```


Renewables, hydro and nuclear {data-icon="fa-signal"}
========================================================================

Column {.tabset .tabset-fade data-width=550 .custom }
-----------------------------------------------------------------------

### __Renewable Energy Generation by source (Terawatt-hours)__ {.no-padding}
```{r fig.height=3}
 
```

### __Contribution by country__ {.no-padding}
```{r fig.height=3}
 
```

### __Hydroelectric generation__ {.no-padding}
```{r fig.height=3}
 
```

### __Nuclear generation__ {.no-padding}
```{r fig.height=3}
 
```


Column {data-width=300}
-----------------------------------------------------------------------

### {.no-padding  }
```{r fig.height=3}

```

Electricity {data-icon="fa-signal"}
========================================================================

Column {.tabset .tabset-fade data-width=550 .custom }
-----------------------------------------------------------------------

### __Electricity generation__ {.no-padding}
```{r fig.height=3}
 
```

### __Contribution by source__ {.no-padding}
```{r fig.height=3}
 

```

### __Share of renewables in power generation__ {.no-padding}
```{r fig.height=3}
 
```


Column {data-width=300}
-----------------------------------------------------------------------

### {.no-padding  }
```{r fig.height=3}

```


About {data-icon="fa-info-circle"}
=======================================================================

Column {data-width=600}
-----------------------------------------------------------------------

### { .custom }

